Automated Topic Naming

### 2012-01-19 ese/ese.tex
minor tweak in intro
Updated clustering

Cleland-Huang and her colleagues have investigated mining requirements
documents for non-functional requirements (NFR) (software
qualities)~\cite{Cleland-Huang2006}.  
%Their approach is similar to
%ours, as they mine keywords from NFR catalogues.  They differ from our approach because
%they mine requirements documents whereas we mine revisions.
One approach they tried was similar to this one, with keywords mined from NFR catalogues found in Chung et al.~\cite{chung99}. 
Their approach resulted in a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
Our research is different because we use a more comprehensive set of terms based on a taxonomy that is an integral part of our framework.
Another difference is that we make cross-project comparisons instead of focusing
on a single project.
They relied on relatively well-structured requirements documents instead of version control histories that we use.
The objective of Cleland-Huang's study was to identify new NFRs for system development, 
while our objective was to recover those latent NFRs from commit-log messages of the project.

@@

% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
In Baldi et al.~\cite{Baldi2008}, topics are named manually: human
experts read the highest-frequency members of a topic and assign a
label accordingly. 
%XXX Possible to cut this
As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
The labels are reasonable enough, but still require an expert in the field to determine them. 
Furthermore, these labels are project specific, because they are
generated from the data of that project. For example we might have a label called \emph{`Oracle'}
in the MySQL case, since Oracle owns MySQL. 
%We do this differently.
%This paper does this differently. 
Our approach differs in two important ways: we automate the process of naming the topics, and we label topics with project-independent terms, in order
to permit cross-project comparison.

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, to control for differences in functional, rather than non-functional, requirements. 
We used three different open-source, partially-commercial database systems:
\begin{description}
\item [ MySQL 3.23] --- Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code~\footnote{generated
using David A. Wheeler's \emph{SLOCCount},
{http://dwheeler.com/sloccount}.}. We used the MySQL 3.23 source control  history from July
31st, 2000 to August 9th, 2004.
\item [MaxDB 7.500] --- Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
has over $940,000$ lines of C source
code~\footnote{{http://www.sdn.sap.com/irj/sdn/maxdb}}. 
 We used the MaxDB 7.500 source control history from June
29th, 2004 to June 19th, 2006.
\item [PostgreSQL 7.3] --- Started in the 1980s as a Berkeley research
  project~\footnote{{http://www.postgresql.org/docs/7.3/static/}}. PostgreSQL
  7.3 contains $306,000$ lines of C code.
 We used the PostgreSQL 7.3 source control  history from May 9th, 2002 to 
 August 26th, 2004.

\end{description}

%Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily
maintenance activities in our studies. We felt that a single domain
would allow for cross-project comparison. At the same time we
recognize that problem domain alone does not guarantee functional and
non-functional similarity.
For instance, each database system has a different focus. PostgreSQL
tends to focus on fulfilling much of the SQL92 specification and adding
more features, while MySQL has been slow to adopt much of that specification in favour of community-requested features.
The consequence of our choice to look at a
single domain is to limit generalizability. What we show in this paper
will be somewhat biased towards database software.

For each project, we studied source control commit comments, the messages
that programmers write when they commit revisions to a source control
repository. 
Most commits we observed had comments: 90\% in MySQL 3.23,
98.5\% in PostgreSQL and 99.99\% in MaxDB 7.500.
Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often
required to create them by the repository mechanism (e.g., Git).  Additionally, relying only on commit comments makes our approach more generalizable,
as we do not assume the presence of other artifact corpora.
%This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such
%message. Our approach is more generalizable if it does not assume the presence of other corpora. 
%bug trackers and email-list archives were not available for both projects. 
An example of a typical commit message, from MySQL, is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were
two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. 
We extracted these messages and indexed them by creation time. 
Each word in the message was
stripped of punctuation and converted to lowercase.
We summarized each message as a word distribution minus stop-words
such as ``\emph{the}'' and ``\emph{at}''. We did not stem or apply any
other transforms to the messages.
Our stop words are derived from the Natural Language Toolkit (NLTK) stop-word English stop-word
list~\footnote{NLTK: \url{http://www.nltk.org/}}.

@@

For the commit message data-sets of each project, we created an XML file that partitioned commits into 30 day periods. 
%This size of period, 30 days, is smaller than the time between minor
%releases but large enough for there to be sufficient commits to
%analyze
We chose a period size of 30 days as it is smaller than the time between minor releases but large enough for there to be sufficient commits to
analyze~\cite{Hindle09ICSM}. 
For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis
algorithm~\cite{Blei2003}, and recorded the topics the algorithm extracted.
%For each project and for each period of messages a project, we input these messages 
%we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.

A topic analysis tool such as LDA will try to find $N$ independent
word distributions within the word distributions of all input
messages. 
If there are not $N$ independent word distributions, the topics
produced tend to be duplicates of each other, that is, they share the
top terms. During this study we found that if we used a value of 20 for $N$, duplicate topics were infrequent.
Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. In other words, these topics are cross-cutting collections of words relevant to one or more of our commit messages. 
LDA extracts topics in an unsupervised manner; the algorithm relies
solely on the source data and word distributions of messages, with no human intervention.
%We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period.

% In topic analysis a single document, such as a commit message, can be related to multiple topics. 
% Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  
% For this paper, a topic represents a word distribution generated from a group of commit log comments which are related by their content.  
%In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

%We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while
%any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}. 
% Furthermore, more than $20$ topics quickly became infeasible for inspection.

@@

To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}, which describes six high-level NFRs: \emph{maintainability, functionality,
portability, efficiency, usability,} and \emph{reliability}. 
%\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
%seems like a reasonable starting point for our work.}.
There is some debate about the terms in this model \cite{Boegh2008}, and whether they are a) the correct terms and b) correctly organized. 
However, ISO9126 is ``an international standard and thus provides an 
internationally accepted terminology for software quality
\cite[p. 58]{Boegh2008},'' and so we consider that it is sufficient for the purposes of this
research.  
\emph{Performance} is an example of an RDBMS word related to the \emph{efficiency} NFR.
We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well suited for comparisons between
projects.

@@

\subsubsection{Creating a Validation Corpus}
To evaluate both semi-unsupervised and supervised classification, we
created a validation set of manually labelled topics. For each project,
 the annotators (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations
%NEIL --> not really! also makes IRR confusing., but some brief
%inspection of annotations was used to confirm that the annotators were acting similarly. 
PostgreSQL (PgSQL) was annotated by both annotators in order
to evaluate inter-rater reliability.
We looked at each period's topics, and assessed what the data---consisting of the frequency-weighted word lists and messages---suggested was the most appropriate labels for that topic. 
We selected the appropriate labels using auxiliary information as well, such as the actual revisions and files that were related to the
topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label. We added a catch-all label, \emph{none}, which we used when none of the six NFRs was suitable. In some instances, we used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?

We validate classification performance using the area under the curve
of Receiver Operating
Characteristic~\cite{Fawcett2006861},
abbreviated \emph{ROC}, sometimes called Area Under the Curve or \emph{AUC}, and the
F-measure, which is the harmonic mean of precision and recall, i.e.,
$2 * (P * R) / (P + R)$. 
Throughout the paper we will provide F-measure scores so that readers
who are more familiar with F-measure than ROC can intuitively
interpret the results.

@@

We argue for using ROC over F-Measure because ROC suffers less from
bias, as F-Measure often skews towards the positive
class~\cite{flach-icml03,Forman:2010:ACS:1882471.1882479}, especially
in the case of class imbalance. Forman et
al.~\cite{Forman:2010:ACS:1882471.1882479} demonstrate that cross-fold
validation will often produce low F-measures given high class
imbalance and the presence of false negatives.
Although
recent work~\cite{Forman:2010:ACS:1882471.1882479} has suggested that while ROC suffers less bias than
the average F-measure ($F_{avg}$) for cross-folds, F-measures ($F_{tp,fp}$) that are calculated from
the sums of true-positives,
false-positive, and false-negatives across folds tend to exhibit no
bias. Unfortunately our experimental framework lacks a robust way to
calculate $F_{tp,fp}$, thus we still provide ROC and $F_{avg}$.
We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).

@@

In order to automatically label each topic with 
one of the six high-level NFRs,
%an NFR from ISO9126, 
we associate each NFR with a list
of keywords or \emph{word-lists}, in our parlance. These word-lists were determined a priori and were not extracted from the projects themselves, using the methodology explained below. In general, these lists are project and domain independent.
We intersected the words of the topics and the words of our word-lists.
We ``labelled'' a topic if any of its words matched any of the word-list's words.
A topic could match more than one NFR.
We used several different sets of word-lists for comparison, which we
refer to as \textsf{exp1}, \textsf{exp2}, and \textsf{exp3} in the text which follows.

Our first word-list set, \textsf{exp1}, was generated using the ontology for software quality measurement described in Kayed et al.~\cite{5072519}, which was constructed using 80 source documents, including research papers and international standards. 
The labels we used were:

\vspace{2mm}
\parbox{0.7\textwidth}{
\small \textsf{
integrity, security,
interoperability, testability, maintainability, traceability,
accuracy, modifiability, understandability, availability, modularity,
usability, correctness, performance, verifiability, efficiency,
portability, flexibility, reliability.
}}
\vspace{2mm}

Our second word-list, \textsf{exp2}, uses the ISO9126 taxonomy described above (Section \ref{sec:wordlist}) to seed the word-lists.
%We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
The terms from ISO9126 may not capture all words occurring in the topics that are nonetheless associated with one of the NFRs. 
For example, the term ``redundancy'' is one 
we considered to be
relevant to discussion of \emph{reliability}, but is not in the
standard. We recognize that terms like this might be used in a
different context with a different meaning, like code-cloning.
We therefore took the NFRs from the ISO9126 and added terms to them.

To construct these expanded word-lists, we used
WordNet~\cite{Fellbaum1998}.
We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `\emph{ilities}' into their respective ISO9126 NFRs. 
We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
We then conducted a simple random analysis of mailing list messages from an open source ecosystem, KDE. Like the GNOME study we conducted in Ernst and Mylopoulos \cite{ernst10refsq}, KDE contains a suite of different products covering a variety of software categories. If we judged a given message to contain terms that were
related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
%Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
% Less info is probably better here...
%For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.

For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using WordNet similarity matches. 
Similarity in WordNet means siblings in a hypernym tree. 
We do not include these words here for space
considerations~\footnote{For our data repository visit \url{http://softwareprocess.es/nomen/}}. 
Wordnet similarity is a very broad match. For example, the label \emph{maintainability} is associated with
words \emph{ease} and \emph{ownership}, and the word \emph{performance} has a `sense' that refers to musical performances, which is obviously unrelated to software development. In general, as we proceed from word-lists in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.

@@

@@

%\DONE{this discussion is about which DB? where do the numbers come from?}
For \textsf{exp1} the labels with the most topics
were
%XXX Neil                    V changed from 748 is that right?
\emph{correctness} (182/305/640, which represent MySQL, MaxDB and PostgreSQL topic counts, respectively) and \emph{testability} (121/238/625). 
We did not see many results for \emph{usability} (4/0/138) or
\emph{accuracy} (3/0/27), which were infrequently matched. Note the significantly higher result for \emph{usability} in PostgreSQL---this suggests a difference in how project developers are discussing usability, at least with respect to our analysis.
We also looked for correlations between our labels: excluding double
matches (self-correlation), our highest co-occurring terms were
\emph{verifiability} or \emph{correctness}
with \emph{traceability}, and \emph{testability} with \emph{correctness}.% (76 and 62 matches, respectively).
% See the following sections for our error analysis.

@@

For \textsf{exp3}, we generally labelled more topics. 
As mentioned above, the word-lists are broad, so there are likely to be false-positives (discussed below). 
The most frequent label across all projects (for this wordlist, and unlike \textsf{exp1}) was \emph{usability}, and the
least frequent label was \emph{maintainability}. This implies that our signifiers for usability in this experiment were fairly broad.
Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability}.
%(200, 190, and 150 topics in common, respectively).

@@

To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine
learners, where the original data is partitioned randomly into ten sub-samples, and each sample is used to test against a training set composed of the nine other samples.
%One subsample is used as a training set and evaluated on the other nine sub-samples. 
%This is repeated nine more times, with each subsample used once as the training set. 
We discuss these results below.% in Section \ref{sec:suplabelling}.

\subsubsection{Analysis of the Supervised Labelling}
% \label{sec:suplabelling}
Because our data-set consisted of word counts we expected Bayesian techniques to perform well. Bayesian techniques are often used in spam filtering, which is similarly interested in word distributions. 
We tried other learners that WEKA~\cite{weka09} provides, including rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best
performing learner per label: 
% We considered the best learner for a label to be 
the learner that had the highest ROC value for that label. 
The best learner is important because one uses a single learner per
label. When applying our technique, for each NFR one should select the
best performing learner possible. Given the tractability of these techniques, a tool which applied all learners and presented the best result should be feasible. 
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

@@

Although we used a variety of machine learners, we found that Bayesian techniques performed the best on all projects. We believe this is due to the large number of features they can handle.
Our best-performing learners---Discriminative Multinomial Naive Bayes, Naive Bayes
and Multinomial Naive Bayes---are all based on Bayes' theorem and all make the naive assumption
that the features supplied are independent. 
%The features we used are word counts per message. 
One beneficial aspect of this result is that it suggests we can have
very fast training and classifying  since training on or classifying one
instance with Naive Bayes can be calculated in $O(N)$
for $N$ features.
%http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html

@@

In Figure \ref{fig:authorcluster} we clustered the developers from
PostgreSQL using the
Ward method of Hierarchical clustering on the 7-d space.  
This will group developers according to a distance metric. 
We clustered the authors into both $2$-clusters
and $6$-clusters. We chose $6$-clusters because the difference in distance
between entities and centroids was minimal at $6$-clusters, and the
derivative of distance was the highest in that region as well.

>>>>

Cleland-Huang and her colleagues have investigated mining requirements
documents for non-functional requirements (NFR) (software
qualities)~\cite{Cleland-Huang2006}.  
%Their approach is similar to
%ours, as they mine keywords from NFR catalogues.  They differ from our approach because
%they mine requirements documents whereas we mine revisions.
One approach they tried was similar to this one, with keywords mined from NFR catalogues found in Chung et al.~\cite{corollary99}. 
Their approach resulted in a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
Our research is different because we use a more comprehensive set of terms based on a taxonomy that is an integral part of our framework.
Another difference is that we make cross-project comparisons instead of focusing
on a single project.
They relied on relatively well-structured requirements documents instead of version control histories that we use.
The objective of Cleland-Huang's study was to identify new NFRs for system development, 
while our objective was to recover those latent NFRs from commit-log messages of the project.

@@

% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
In Baldi et al.~\cite{Baldi2008}, topics are named manually: human
experts read the highest-frequency members of a topic and assign a
label accordingly. 
%XXX Possible to cut this
As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
The labels are reasonable enough, but still require an expert in the field to determine them. 
Furthermore, these labels are project-specific, because they are
generated from the data of that project. For example we might have a label called \emph{`Oracle'}
in the MySQL case, since Oracle owns MySQL. 
%We do this differently.
%This paper does this differently. 
Our approach differs: first of all, we automate the process of naming the topics; secondly, we label topics with project-independent terms, in order
to permit cross-project comparison.

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, to control for differences in functional, rather than non-functional, requirements. 
We used three different open-source, partially-commercial database systems:
\begin{itemize}
\item  MySQL 3.23. Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code~\footnote{generated
using David A. Wheeler's \emph{SLOCCount},
{http://dwheeler.com/sloccount}.}. We used the MySQL 3.23 source control  history from July
31st, 2000 to August 9th, 2004.
\item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project
has over $940,000$ lines of C source
code~\footnote{{http://www.sdn.sap.com/irj/sdn/maxdb}}. 
 We used the MaxDB 7.500 source control history from June
29th, 2004 to June 19th, 2006.
\item PostgreSQL 7.3. Started in the 1980s as a Berkeley research
  project~\footnote{{http://www.postgresql.org/docs/7.3/static/}}. PostgreSQL
  7.3 contains $306,000$ lines of C code.
 We used the PostgreSQL 7.3 source control  history from May 9th, 2002 to 
 August 26th, 2004.

%Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily
maintenance activities in our studies. We felt that a single domain
would allow for cross-project comparison. At the same time we
recognize that a problem-domain alone does guarantee functional and
non-functional similarity.
For instance each database system has a different focus, PostgreSQL
tends to focus on fulfilling much of the SQL92 specification and adding
more features while MySQL has been slow to adopt much of the spec.
 As a consequence our choice of looking at a
single domain limits our generalizability. What we show in this paper
will be somewhat biased towards database software.

For each project, we used source control commit comments, the messages
that programmers write when they commit revisions to a source control
repository. 
Most commits we observed had commit comments (90\% in MySQL 3.23,
98.5\% in PostgreSQL and 99.99\% in MaxDB 7.500).
Commit comments are often studied by researchers, as they are the most readily accessible source of project interactions, and developers are often
required to create them by the repository mechanism (e.g., Git).  Additionally, relying only on commit comments makes our approach more generalizable,
as we do not assume the presence of other artifact corpora.
%This is the most readily accessible source of project interactions for outside researchers, since developers often or always (e.g., Git) write such
%message. Our approach is more generalizable if it does not assume the presence of other corpora. 
%bug trackers and email-list archives were not available for both projects. 
An example of a typical commit message, from MySQL, is: \textit{``history annotate diffs bug fixed (if mysql\-\_real\-\_connect() failed there were
two pointers to malloc'ed strings, with memory corruption on free(), of course)''}. 
We extracted these messages and indexed them by creation time. 
Each word in the message was
stripped of punctuation and converted to lowercase.
We summarized each message as a word distribution minus stop-words
such as ``\emph{the}'' and ``\emph{at}''. We did not stem or apply any
other transforms to the messages.
Our stop words are derived from the Natural Language Toolkit (NLTK) stop-word English stop-word
list~\footnote{NLTK: \url{http://www.nltk.org/}}.

@@

For the commit message data-sets of each project, we created an XML file that separated commits into 30 day periods. 
%This size of period, 30 days, is smaller than the time between minor
%releases but large enough for there to be sufficient commits to
%analyze
We chose a period size of 30 days as it is smaller than the time between minor releases but large enough for there to be sufficient commits to
analyze~\cite{Hindle09ICSM}. 
For each 30 day period of each project, we input the messages of that period into Latent Dirichlet Allocation (LDA), a topic analysis
algorithm~\cite{Blei2003}, and recorded the topics the algorithm extracted.
%For each project and for each period of messages a project, we input these messages 
%we created `topics' using Latent Dirichlet Allocation, a topic identification algorithm.

A topic analysis tool such as LDA will try to find $N$ independent
word distributions within the word distributions of all input
messages. 
If there are not $N$ independent word distributions, the topics
produce tend to be duplicates of each other, that is they share the
top terms. During this study we found that when $N$ was 20 we did not
witness duplicate topics very often.
Linear combinations of these $N$ word distributions are meant to represent and recreate the word distributions of any of the original messages. 
These $N$ word distributions effectively form \emph{topics}: cross-cutting collections of words relevant to one or more of our commit messages. 
LDA extracts topics in an unsupervised manner; the algorithm relies
solely on the source data and word distributions of messages, with no human intervention.

In topic analysis a single document, such as a commit message, can be related to multiple topics. 
Representing documents as a mixture of topics maps well to source code repository commits, which often have more than one purpose~\cite{Hindle09ICSM}.  
For this paper, a topic represents a word distribution generated from a group of commit log comments which are related by their content.  
%In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

We applied Blei's LDA implementation~\cite{Blei2003} against the word distributions of these commits, and generated lists of topics per period. 
We set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while
any more topics dilutes the results and creates indistinct topics~\cite{Hindle09ICSM}. 
% Furthermore, more than $20$ topics quickly became infeasible for inspection.

@@

To facilitate cross-project comparison, we used a taxonomy of NFRs. This taxonomy is based on the ISO quality model, ISO9126~\cite{iso9126}. 
ISO9126 describes six high-level NFRs: \emph{maintainability, functionality,
portability, efficiency, usability,} and \emph{reliability}
%\footnote{While there may be lingering debate in some circles about these terms, an ISO standard
%seems like a reasonable starting point for our work.}.
There is some debate about the terms in this model, and whether they are a) the correct terms and b) correctly organized. 
However, ISO9126 is ``an international standard and thus provides an 
internationally accepted terminology for software quality
\cite[p. 58]{Boegh2008},'' that is sufficient for the purposes of this
research.  
\emph{Performance} is an example of RDBMS word related to the \emph{efficiency} NFR.
We claim that these NFRs are maintenance concerns (to varying degrees) in all software projects, and are therefore well-suited for comparisons between
projects.

@@

\subsubsection{Creating a Validation Corpus}
To evaluate both semi-unsupervised and supervised classification, we
created a validation set of manually labelled topics. For each project,
 the annotators (the first two authors) annotated each extracted topic in each period with
the six NFR labels listed above.
Annotators did not annotate each other's annotations, but some brief
inspection of annotations was used to confirm that the annotators were
acting similarly. PostgreSQL (PgSQL) was annotated by both annotators in order
to evaluate inter-rater reliability.
We looked at each period's topics, and assessed what the data ---
consisting of the frequency-weighted word lists and messages ---
suggested was the label for that topic. 
We were able to pinpoint the appropriate label using auxiliary information as well, such as the actual revisions and files that were related to the
topic being annotated.
For example, for the MaxDB topic consisting of a message ``exit() only
used in non NPTL LINUX Versions'', we tagged that topic
\emph{portability}. 
Given the top-level annotations of \emph{none}, \emph{portability},
\emph{efficiency}, \emph{reliability}, \emph{functionality},
\emph{usability}, and \emph{maintainability}, the annotators annotated each topic
with the relevant label. Sometimes they used finer-grained
annotations that would be aggregated up to one of these higher-level labels.
%We compared against this data-set, but we also used the data-set for our supervised machine learning based topic classification. 
% A topic may have more than one matching keyword. % ok what?

We validate classification performance using the area under the curve
of Receiver Operating
Characteristic~\cite{Fawcett2006861},
abbreviated \emph{ROC}, sometimes called \emph{AUC}, and the
F-measure, which is the harmonic mean of precision and recall, i.e.,
$2 * (P * R) / (P + R)$. 
Throughout the paper we will provide F-measure scores so that readers
who are more familiar with F-measure than ROC can intuitively
interpret the results.

@@

We argue for using ROC over F-Measure because ROC suffers less from
bias than F-Measure and F-Measure often skews towards the positive
class~\cite{flach-icml03,Forman:2010:ACS:1882471.1882479}, especially
in the case of class imbalance. Forman et
al.~\cite{Forman:2010:ACS:1882471.1882479} demonstrate that cross-fold
validation will often produce low F-measures given high class
imbalance and the presence of false negatives.
Although
recent work~\cite{Forman:2010:ACS:1882471.1882479} has suggested that while ROC suffers less bias than
the average F-measure ($F_{avg}$) for cross-folds, F-measures ($F_{tp,fp}$) that are calculated from
the sums of true-positives,
false-positive, and false-negatives across folds tend to exhibit no
bias. Unfortunately our experimental framework lacks a robust way to
calculate $F_{tp,fp}$, thus we still provide ROC and $F_{avg}$.
We consider our labelling classifiers acceptable if they outperform a random classifier (0.5).

@@

In order to automatically label each topic with 
one of the six high-level NFRs,
%an NFR from ISO9126, 
we associate each NFR with a list
of keywords (\emph{word-lists}, in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves, using the methodology explained below. In general, these lists are project and domain independent.
We intersected the words of the topics and the words of our word-lists.
We ``labelled'' a topic if any of its words matched any of the word-list's words.
A topic could match more than one NFR.
We used several different sets of word-lists for comparison, which we
refer to as \textsf{exp1}, \textsf{exp2}, and \textsf{exp3} in the text which follows.

Our first word-list set, \textsf{exp1}, was generated using the ontology described in Kayed et al.~\cite{5072519}
That paper constructs an ontology for software quality measurement using eighty source documents, including research papers and international standards. 
The labels we used are:

\begin{quotation}
\small \noindent \textsf{
integrity, security,
interoperability, testability, maintainability, traceability,
accuracy, modifiability, understandability, availability, modularity,
usability, correctness, performance, verifiability, efficiency,
portability, flexibility, reliability.
}
\end{quotation}

Our second word-list, \textsf{exp2}, uses the ISO9126 taxonomy described above (Section \ref{sec:wordlist}) to seed the word-lists.
%We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
The terms from ISO9126 may not capture all words occurring in the topics that are nonetheless associated with one of the NFRs. 
For example, the term ``redundancy'' is one 
we considered to be
relevant to discussion of \emph{reliability}, but is not in the
standard. We recognize though that terms like this might be used in a
different context with a different meaning, like code-cloning.
We therefore took the NFRs from the ISO9126 and added terms to them.

To construct these expanded word-lists, we used
WordNet~\cite{Fellbaum1998}.
We then added Boehm's software quality model~\cite{Boehm+:1976:ICSE}, and classified his eleven `\emph{ilities}' into their respective ISO9126 NFRs. 
We did the same for the quality model produced by McCall et al.~\cite{mccall1977}. 
We then did a simple random analysis of mailing list messages from an open source ecosystem, KDE. Like the GNOME study we conducted in Ernst and Mylopoulos \cite{ernst10refsq}, KDE contains a suite of different products covering a variety of software categories. If we judged a given message to contain terms that were
related to one of the NFRs in ISO9126, we added it to our word-list. This allowed us to expand our word-lists with more software-specific terms.
%Finally, we randomly analyzed two mailing lists from another software project to expand our set with domain-specific terms. 
% Less info is probably better here...
%For example, we add the term ``performance'' to the synonyms for \emph{efficiency}, since this term occurs in most mail messages that discuss efficiency.
Table \ref{tbl:wnsig} shows the labels (NFRs) and word-lists we used for matching.

For the third set of word-lists, \textsf{exp3}, we extended the word-lists from \textsf{exp2} using WordNet similarity matches. 
Similarity in WordNet means siblings in a hypernym tree. 
We do not include these words here for space
considerations~\footnote{For our data repository visit \url{http://softwareprocess.es/nomen/}}. 
Wordnet similarity is a very broad match. For example, the label \emph{maintainability} is associated with
words \emph{ease} and \emph{ownership}, and the word \emph{performance} has a `sense' which refers to musical performances -- which is not related to software development. In general, as we proceed from word-lists in \textsf{exp1} to that in \textsf{exp3}, our lists become more generic.

@@

%\noindent \textbf{Generating labels automatically} ---

@@

%\DONE{this discussion is about which DB? where do the numbers come from?}
For \textsf{exp1} the labels with the most topics
were
%XXX Neil                    V changed from 748 is that right?
\emph{correctness} (182/305/640, which represent MySQL, MaxDB and PostgreSQL topic counts, respectively) and \emph{testability} (121/238/625). 
We did not see many results for \emph{usability} (4/0/138) or
\emph{accuracy} (3/0/27), which were infrequently matched. Note the significantly higher result for \emph{usability} in PostgreSQL -- this suggests a difference in how this project is discussing usability, at least with respect to our analysis.
We also looked for correlations between our labels: excluding double
matches (self-correlation), our highest co-occurring terms were
\emph{verifiability} or \emph{correctness}
with \emph{traceability}, and \emph{testability} with \emph{correctness}.% (76 and 62 matches, respectively).
% See the following sections for our error analysis.

@@

For \textsf{exp3}, we generally labelled more topics. 
As we mentioned, the word-lists are broad, so there are likely to be false-positives (discussed below). 
The most frequent label across all projects (for this wordlist, and unlike \textsf{exp1}) was \emph{usability}, and the
least frequent label was \emph{maintainability}. This implies that our signifiers for usability in this experiment were fairly broad.
Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability}.
%(200, 190, and 150 topics in common, respectively).

@@

To assess the performance of the supervised learners, we did a 10-fold cross-validation~\cite{Kohavi1995}, a common technique for evaluating machine
learners. 
The original data is partitioned randomly into ten sub-samples. Each sample is used to test against a training set composed of the nine other samples.
%One subsample is used as a training set and evaluated on the other nine sub-samples. 
%This is repeated nine more times, with each subsample used once as the training set. 
We have reported these results below.% in Section \ref{sec:suplabelling}.

\subsubsection{Analysis of the Supervised Labelling}
% \label{sec:suplabelling}
Because our data-set was of word counts we expected Bayesian techniques, often used in spam filtering, to perform well. 
We tried other learners that WEKA~\cite{weka09} provides: rule learners, decision tree learners, vector space learners, and support vector machines.  
Figure \ref{fig:best-learn-per-tag} shows the performance of the best
performing learner per label: 
% We considered the best learner for a label to be 
the learner that had the highest ROC value for that label. 
The best learner is important because one uses a single learner per
label. When applying our technique, for each NFR one should select the
best performing learner possible. Given the tractability of these techniques, a tool which applied all learners and presented the best result should be feasible. 
%Figure \ref{fig:best-learn-per-tag} uses the ZeroR learner as a baseline, a common practice in machine learning,
%as ZeroR naively chooses the largest category all of the time. 
%ZeroR occasionally outperforms other learners; for labels which are not as common, this is to be expected because any miscategorization will hurt accuracy. 
%This is why we use ROC values, instead of accuracy, as they can better represent performance on labels which are not applicable to the majority of samples.

@@

For all projects Bayesian techniques did the best out of a wide variety of machine learners tested, due to the large number of features they can handle.
Our best learners, Discriminative Multinomial Naive Bayes, Naive Bayes
and Multinomial Naive Bayes  are all based on Bayes' theorem and all
assume, naively, that the features supplied are independent. 
%The features we used are word counts per message. 
One beneficial aspect of this result is that it suggests we can have
very fast training and classifying  since training on or classifying one
instance with Naive Bayes can be calculated in $O(N)$
for $N$ features.
%http://nlp.stanford.edu/IR-book/html/htmledition/naive-bayes-text-classification-1.html

@@

In Figure \ref{fig:authorcluster} we clustered the developers from
PostgreSQL using the
Ward method of Hierarchical clustering on the 7-dimensional space.  
This will group developers according to the Euclidean distance metric.
%We chose cosine distance because it is often used in IR research to
%compare documents, in this case we are comparing the labels of topics
%associated with each author's commits.
The height in Figure \ref{fig:authorcluster} describes the euclidean
distance between elements, in terms of the path 
We clustered the authors into both $2$-clusters
and $6$-clusters. We chose $6$-clusters because the difference in distance
between entities and centroids was minimal at $6$-clusters, and the
derivative of distance was the highest in that region as well.
